FraCaS: Temporal Analysis

### 2020-03-10 paper2020/paper.tex
revision

- See our pervious papers, (adjectives, adverbs, nouns, verbs, anaphora, etc.)

- Here we simply ignore these aspects and it is safe for the reader to
assume a montegovian semantics and assume that anaphoroi are resolved
as they should be.

@@

Our basic approach is to extend predicates with two time parameters,
one corresponding with the starting time of the action and one with
its stopping time. For example, if John walked between $t_0$ and
$t_1$, we would have: $walk(john,t0,t1)$. From now on we will call an
interval of time points $[t_0,t_1]$ a timespan. Every timespan
$[t_0,t_1]$ has the property $t_0 ≤ t_1$ (it starts no later than it
stops).

@@

We adjust the montegovian semantics so that every category takes a
\emph{optional} timespan as an additional parameter. This timespan
serves as a time reference for semantics, and we call it a
\emph{temporal context}. This context propagates through the
compositional interpretation down to atoms. By default, every
interpretation passes the temporal context down to its components
without changing it. However some key elements will act on it on
nontrivial ways, which we proceed to detail below.

@@

The prime non-trivial manipulation of timespans are tenses. In our
syntax, inherited from GF, tenses are represented syntactically as an
attribute of the clause. TODO{show example}

>>>>

We are relying on a largely standard montegovian semantics. In brief,
sentences are interpreted as propositions, verbs and noun-phrases as
predicates.  We use type-raising for noun-phrases to support
quantifiers.

We support additionally various constructinos and phenomena present in
the testsuite, including adjectives, adverbs, nouns, verbs, anaphora,
etc. We direct the reader to our previous work for details \ifanon
(citations anonymized) \else \citet{} \fi, but the particular
treatment of such phenomena is largely independent from our treatment
of time.
%
Here we simply ignore these aspects beyond the fact that they
are handled correctly in most of the case of the FraCaS testsuite.  In
particular, non-temporal anaphoroi are resolved as they should be.

@@

In montegovian semantics, (intransitive) verbs are one-place
predicates; in types, they are functions from entities to propositions
($e \to t$).  Our basic approach is to generalise verb to take two
additional time parameters, one corresponding with the starting time
of the action and one corresponding to its stopping time. For example,
if John walked between $t_0$ and $t_1$, we would have:
$walk(john,t0,t1)$. From now on we will call an interval of time
points $[t_0,t_1]$ a timespan. Every timespan $[t_0,t_1]$ has the
property $t_0 ≤ t_1$ (it starts no later than it stops). We are thus
using a simple newtonian model of time, corresponding to layman
intuition.

In principle, common nouns and adjectives should undergo the same
procedure. For simplicity we will however only consider verbs from now
on. (In fact in our implementation we chose not to modify nouns nor
adjectives to limit the increase in complexity of the formulas.)

@@

We adjust the montegovian semantics so that the interpretation of
every category (propositions, verb phrases, etc.) takes an
\emph{optional} timespan as an additional parameter. This timespan
serves as a time reference for semantics, and we call it a
\emph{temporal context}. This context propagates through the
compositional interpretation down to lexical items with atomic
representation (verbs). By default, every interpretation passes the
temporal context down to its components without changing it. However
some key elements will act on it on nontrivial ways, which we proceed
to detail below.

@@

The principal non-trivial manipulation of timespans are tense
markers. In our syntax, inherited from GF, tenses are represented
syntactically as an attribute of the clause.

Example:

A scandinavian won the nobel prize.

useCl past pPos (predVP (detCN (detQuant indefArt numSg) (useN (lexemeN "scandinavian_N"))) (complSlash (slashV2a (lexemeV2 "win_V2")) (detCN (detQuant indefArt numSg) (useN (lexemeN "nobel_prize_N")))))
